---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
library(countries)
```

# countries 

<!-- badges: start -->
<!-- badges: end -->

`countries` is an R package designed to quickly wrangle, merge and explore country data. This package contains functions to easily identify and convert country names, pull country info and datasets, merge country data from different sources, and easily make world maps.

## Installing and loading the package

The package can now be installed from CRAN.

```{r , eval=FALSE}
# Install package from CRAN
install.packages("countries")

# load package
library(countries)
```


## Dealing with country names

The package contains several functions to work with country names. For instance, the function `is_country()` can be used to test for country names or subsets of countries: 

```{r}
#Detect strings that are country names
is_country(x=c("ITA","Estados Unidos","bungalow","dog",542))

#Checking for a specific subset of countries
is_country(x=c("Ceylon","LKA","Indonesia","Inde"), check_for=c("India","Sri Lanka"))
```

The function `country_name()` can be used to convert country names to different naming conventions or to translate them to different languages. `country_name()` can identify countries even when they are provided in mixed formats or in different languages. It is robust to small misspellings and recognises many alternative country names and old nomenclatures.  

```{r}
example <- c("US","C@ète d^Ivoire", "Morocco","FYROM", "Arabie Saoudite")

# Getting 3-letters ISO code
country_name(x= example, to="ISO3")

# Translating to spanish
country_name(x= example, to="name_es")

# Getting multiple nomenclatures
country_name(x= example, to=c("ISO3","ISO2","UN_en"))
```

Learn more about how to deal with country names in [this article](https://fbellelli.github.io/countries/articles/dealing_with_names.html).


## Getting information about countries

`country_info()` allows to download a variety of information about countries, such as: currencies used, capital city, language spoken, flag, neighbouring countries, and much more.

```{r}
# What are the official languages of Switzerland?
country_info("Switzerland", "languages")

# Get information on the capital name and currencies for multiple countries
country_info(c("Canada", "Mozambique", "India"), c("capital", "currencies"))
```

You can find [more information about this function](https://fbellelli.github.io/countries/reference/country_info.html) in the documentation.


## Easy country maps

With `quick_map()`, it takes only one line of code to produce [chloropleth maps](https://en.wikipedia.org/wiki/Choropleth_map). It automatically recognises country names in multiple languages and nomenclatures. This allows to produce publication-grade map in seconds. Moreover, the output is a ggplot object, so the visual look can be customised in infinite ways.

```{r, dpi = 700, warning = FALSE, message = FALSE, fig.asp=0.5}
# downloading some sample data to plot
example_data <- country_info(fields = c("car"))

# make a map
quick_map(example_data, plot_col = "car.side")
```

You can find [more examples in this article](https://fbellelli.github.io/countries/articles/quick_map.html).

## Merging country data

The function `auto_merge()` simplifies the merging of country data tables by: **1)** allowing merging of 2+ tables at the same time, **2)** Supporting automatic detection of columns to merge, **3)** automatically handling different country naming conventions and date formats, **4)** automatic pivoting of country names or years in tables' headers.

```{r}
# Let's create 4 tables with different formats and country names
tab1 <- data.frame(country = c("Italy", "Pakistan", "Brazil"), world_cups = c(4, 0, 5))
tab2 <- data.frame(exporter = c("DEU", "DEU", "ITA", "ITA"), HS_chapter = c(9, 85, 9, 85), volume = c(800, 5000, 1000, 2000))
tab3 <- data.frame(HS = c(9, 85), Description = c("Coffee, tea and mate", "Electrical machinery"))
tab4 <- data.frame(year = c(2010, 2011), Allemagne = runif(2), Brésil = runif(2), Pakistan = runif(2))

# These tables can easily be merged with one line of code:
auto_merge(tab1, tab2, tab3, tab4)

```


Learn more about country names functions in [this article](https://fbellelli.github.io/countries/articles/auto_merge.html).


