---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
```

# countries 

<!-- badges: start -->
<!-- badges: end -->

`countries` is an R package designed to quickly wrangle, merge and explore country data. This package contains functions to easily identify and convert country names, pull country info and datasets, merge country data from different sources, and make quick world maps.

## Installing and loading the package

Since the package is not yet on CRAN, the development version needs to be downloaded directly from the Github repository. This can be done with the `devtools` package.

```{r , eval=FALSE}
# Install and load devtools
install.packages("devtools")
library(devtools)

# Install countries
devtools::install_github("fbellelli/countries", build_vignettes = TRUE)
```

The package can then be loaded normally

```{r}
library(countries)
```


## Dealing with country names

The package contains several functions to work with country names. For instance, the function `is_country()` can be used to test for country names or subsets of countries: 

```{r}
#Detect strings that are country names
is_country(x=c("ITA","Estados Unidos","bungalow","dog",542))

#Checking for a specific subset of countries
is_country(x=c("Ceylon","LKA","Indonesia","Inde"), check_for=c("India","Sri Lanka"))
```

The function `country_name()` can be used to convert country names to different naming conventions or to translate them to different languages. `country_name()` can identify countries even when they are provided in mixed formats or in different languages. It is robust to small misspellings and recognises many alternative country names and old nomenclatures.  

```{r}
example <- c("US","C@ète d^Ivoire", "Morocco","FYROM", "Arabie Saoudite")

# Getting 3-letters ISO code
country_name(x= example, to="ISO3")

# Translating to spanish
country_name(x= example, to="name_es")

# Getting multiple nomenclatures
country_name(x= example, to=c("ISO3","ISO2","UN_en"))
```

Learn more about country names functions in [this article](/articles/dealing_with_names.html).

## Work in progress:

* function for downloading up-to-date information on countries (e.g. currency, language, population, etc.)
* function for downloading country data for analysis from different sources (e.g. UN, World Bank, FRED, etc.)
* function to quickly merge country data from different sources
* function to tag countries based on common criteria (e.g. developing status, World bank income group, geographic region, etc.) 
* function to easily plot chloropleth maps
* publish on CRAN
