---
title: "Easy country chloropleth maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Easy country chloropleth maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- avoid border around images -->
<style>
    img {
        border: 0;
    }
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[Chloropleth maps](https://en.wikipedia.org/wiki/Choropleth_map) can be a powerful visual tool to communicate information about countries. However, making a nice map in R can be deceivingly time-consuming. This is where `quick_map()` can help! 

Its fuzzy matching algorithm automatically recognises country names and cuts the hassle of matching country data to country shapes. Predefined colour themes and zooms are provided, to cut the time involved in making a nice looking map. It even provides an easy integrated option to save the plot locally with predefined settings.

With `quick_map()` you can produce a publication-grade plot with one short line of code. If you are still not happy with how your map looks, customisation is endless with all the usual ggplot commands. 

In this vignette, we will walk through some examples.

## Some data to plot

Let's start by loading the package and putting together a table with country data which we will use for demonstration. This table contains information on countries' population (numeric continuous variable), a boolean variable indicating whether the country is landlocked, a categorical/factor variable indicating whether cars in the country drive on the left or right, and finally information on the capital of the countries (name, latitude/longitude).

```{r}
library(countries)

data <- country_info(fields = c("population", "landlocked", "car", "capital", "capitalInfo"))
``` 

## Quick start

All it takes to make a plot is one line of code telling to plot the column `"car.side"`from the table `data`. The function will take care of everything else. Under the hood, the function looks for a column of country names and matches them to one of the territories in the [ISO standard 3166-1](https://en.wikipedia.org/wiki/ISO_3166-1).

**NOTE: The resolution of the images has been has been intentionally reduced in this vignette to save space on CRAN.** 

```{r, dpi = 200, warning = FALSE, message = FALSE, fig.asp = 0.5, fig.width = 7,  out.width="100%"}
quick_map(data, plot_col = "car.side", verbose = FALSE)
``` 

### Saving your plot

The plot can be saved with by passing a file path to the argument `save_to`. This comes with predefined size and resolution options. If you wish a more customised download, the function `ggplot2::ggsave()` can be used.

```{r, eval = FALSE}
quick_map(data, plot_col = "car.side", 
          save_to = "example_path/plot.png")
```


### What happens if a country is not recognised ?

When a country name is not recognised or does not match with a country in the standard, a warning is given to the user. In the example above, the country that has no match (UNK) is Kosovo, which is not in the ISO 3166-1 standard.

If you are interested in knowing exactly how the country were matched you can just use the function `country_name()`.


## Zooming / Regional maps

Suppose you wish to plot a map of specific regions of the globe. The package comes with predefined zoom settings `c("Default", "World", "Africa", "Asia", "Europe", "SEAsia", "NAmerica", "CAmerica", "SAmerica", "Oceania")`. Alternatively, it is also possible to define a custom zoom by passing the min/max longitude and latitude.

```{r, warning = FALSE, message = FALSE, dpi = 200, out.width= "50%",  fig.align = "center"}
quick_map(data, plot_col = "car.side", zoom = "Africa")
``` 


```{r,  warning = FALSE, message = FALSE, dpi = 200, out.width= "50%",  fig.align = "center"}
quick_map(data, plot_col = "car.side", zoom = c(25, 55, -10, 15))
``` 

## Customising the look of your map

The function provides basic settings to customise the essential elements of the map in a simple way. If these changes are not enough, full customisation can be achieved with ggplot commands. Let's see some examples.

### Themes and visual settings

With the argument `theme` it is possible to switch among the predefined looks for the map. `theme` can take as a value a number between 1 and 11, or one of the predefined themes' names: `c("Default", "Greyscale", "Candy", "RedBlue", "Dark", "Reds", "Blues", "Greens", "Viridis", "Cividis", "Distinct")`. If `0` or `"NoTheme"` is passed, ggplot2 settings are used (no theme is applied to the map).

```{r,  dpi = 170, warning = FALSE, message = FALSE, fig.asp = 0.5, fig.width = 7,  out.width="100%"}

quick_map(data, plot_col = "population", verbose = FALSE,
          theme = "Candy")

``` 


Some basic visual elements can be modified within `quick_map()`. For instance, In case you are not happy with the colour ordering, it can be reversed with `reverse_palette = TRUE`. `col_na` can be used to set the colour for countries with missing data (NAs), the variable name can be changed with `name_legend`, finally, `col_na` and `width_border` set the colour and thickness of the border lines between territories (or remove them altogether). 

```{r, dpi = 200, warning = FALSE, message = FALSE, fig.asp = 0.5, fig.width = 7,  out.width="100%"}
quick_map(data[1:150,], plot_col = "population", verbose = FALSE,
          theme = "Greyscale",
          reverse_palette = TRUE,
          name_legend = "Countries' populations\nas of 2022", 
          col_na = "darkred",
          col_border = "NA") # here we are removing the border lines

``` 


### Colour scale breaks


By default, `quick_map()` discretises continuous variables, such as population, in 6 buckets for plotting. This is because continuous scales usually are not easy to interpret in world maps. This means that the colour assignment depends the bucket in which the country falls. In some cases, like in the example above, the automatic buckets defined by `quick_map()` may not yield the best results. Users can supply values for the colour breaks manually with the argument `col_breaks`. In the example below, I manually pass powers of ten as breaks.

```{r, dpi = 200, warning = FALSE, message = FALSE, fig.asp = 0.5, fig.width = 7,  out.width="100%"}
quick_map(data, plot_col = "population", verbose = FALSE,
          theme = "RedBlue",
          col_breaks = c(0, 1e5, 1e6, 1e7 , 1e8, 1e9))

``` 

Some predefined themes have only a limited number of colours. If you have many breaks, the following themes support more colours: `c("Greyscale", "Dark", "Reds", "Blues", "Greens", "Viridis", "Cividis", "Distinct")`. Finally, if you wish to keep a continuous colour scale, instead of a disretised one, you can request the raw ggplot output with `theme = 0`.





### Advanced customisation

The nice thing about `quick_map()` is that the output is a ggplot object, so customisation is truly endless. Let's start with our last plot:

```{r, message=FALSE, warning=FALSE}
p <- quick_map(data, plot_col = "population", verbose = FALSE,
          theme = "RedBlue",
          col_breaks = c(0, 1e5, 1e6, 1e7 , 1e8, 1e9))

``` 

The look of the can be customised with any of ggplot's `theme` or `scale` commands. For example, below I remove the name of the legend, **provide custom colours** for the 6 buckets, add a title and change the background:

```{r,  dpi = 200, warning = FALSE, message = FALSE, fig.asp = 0.5, fig.width = 7,  out.width="100%"}
library(ggplot2)

# let's define 6 colours for our map
colours6 <- c("white", "lightblue", "blue", "grey", "black","yellow")

# let's apply them to our plot!
p +
  scale_fill_manual(name = NULL,  # remove name in fill legend
                    values = colours6)+ # use custom colours
  ggtitle("World population") + # add a title to the plot 
  theme(plot.background = element_rect(fill = "#e3cff9"), # change colour of background
        legend.position = "top",  # move legend on top
        plot.title = element_text(hjust = 0.5, family = "mono")) # center the title and change font family

``` 

Anything that works with ggplot will work also with the output of `quick_map()`. It's even possible to **add annotations or additional layers** to the map. For example, in the image below, I am adding text, arrows and markers corresponding to country capitals by using the longitude and latitude as coordinates.

```{r,  dpi = 200, warning = FALSE, message = FALSE, fig.asp = 0.6,  fig.width = 8, out.width="100%"}
# extract capital longitude and latitude in data table
data$long <- as.numeric(sub(x = data$capitalInfo.latlng, pattern = "^.*; ", replacement = "", perl = TRUE))
data$lat <- as.numeric(sub(x = data$capitalInfo.latlng, pattern = ";.*$", replacement = "", perl = TRUE))

# make a map
quick_map(data, "landlocked", zoom = "Europe", theme = "Greens")+
  geom_point(data = data, mapping = aes(x = long, y = lat, group = NULL), shape = 10) + # add markers for the capitals
  geom_text(data = data[100:120,], mapping = aes(x = long, y = lat, label = capital, group = NULL), size = 3, nudge_y = -1) + # printing capital name for first 20 rows
  annotate("segment", x = 18, y = 35, xend = 14, yend = 42,
           arrow = arrow(type = "closed", length = unit(0.02, "npc")))+   # add an arrow
  annotate("label", x = 18, y = 35, label = "A boot!")  # add a text annotation


``` 

